Gradual Soundness: Lessons from Static Python

### 2022-02-01 pj-7.1/paper.tex
pj: threats -> threat
pj: nbody data

  \centering
  \begin{tabular}{lrr}
    Name               & Typed / Python & Refined / Python \\\midrule
    \bmname{deltablue} &         0.59 &           0.30 \\
    %% \bmname{nbody}     &         FILL &           FILL \\
    \bmname{richards}  &         0.53 &           0.22 \\
    \bmname{fannkuch}  &         1.03 &           0.46
    %% fannkuch typedbasic2 = 0.90
  \end{tabular}

@@

Almost all ratios are less than one, showing that \SP{} achieves a speedup over
Python despite the cost of enforcing soundness.
The one exception is the basic typed version of \bmname{fannkuch}, which sees a minor
slowdown.
This benchmark is primarily about number-crunching and the basic version uses standard
numeric types that \SP{} does not optimize.
Switching to primitive types without otherwise changing the code
gives an immediate speedup (ratio = $0.90$).

@@

\paragraph{Threats to Validity}

@@

\begin{table}[tp]
  \caption{Exact commands to invoke each benchmark (1/2)}
  \label{t:mb:commands}
  \tiny\centering

@@

    Fannkuch \colname{T-Min-2} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_fannkuch\_static\_basic2.txt -X jit-enable-jit-list-wi} \\

@@

    Fannkuch \colname{T-Min-2} & \texttt{time ./python -X install-strict-loader Tools/benchmarks/fannkuch\_static\_basic2.py 5} \\

@@

    Fannkuch \colname{T-Min-2} & \texttt{time ./python Tools/benchmarks/fannkuch\_static\_basic2.py 5} \\

@@

    () &

@@

\begin{table}[tp]
  \caption{Exact commands to invoke each benchmark (2/2)}
  \label{t:mb:commands}
  \tiny\centering

@@

    DeltaBlue \colname{T-Min-2} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_deltablue\_static\_basic2.txt -X jit-enable-jit-list-w} \\

@@

    DeltaBlue \colname{T-Min-2} & \texttt{time ./python -X install-strict-loader Tools/benchmarks/deltablue\_static\_basic2.py 100} \\

@@

    DeltaBlue \colname{T-Min-2} & \texttt{time ./python Tools/benchmarks/deltablue\_static\_basic2.py 100} \\

@@

>>>>

  \centering
  \begin{tabular}{lrr}
    Name               & Typed / Python & Refined / Python \\\midrule
    \bmname{deltablue} &         0.59 &           0.30 \\
    \bmname{fannkuch}  &         1.03 &           0.46 \\
    \bmname{nbody}     &         1.09 &           0.24 \\
    \bmname{richards}  &         0.53 &           0.22
    %% fannkuch typedbasic2 = 0.90
  \end{tabular}

@@

Almost all ratios are less than one, showing that \SP{} achieves a speedup over
Python despite the cost of enforcing soundness.
The exceptions are the simply-typed version of \bmname{fannkuch} and
\bmname{nbody}.
Both these benchmarks focus on numeric operations, which \SP{} does not
optimize because it offers primitive types.
Indeed, changing \bmname{fannkuch} to use a few primitive types without
otherwise changing the code gives an immediate speedup (ratio = $0.90$).

@@

\paragraph{Threat to Validity}

@@

\begin{table}[tp]
  \caption{Nbody microbenchmark data}
  \label{t:mb:nbody}
  \footnotesize\centering

  \begin{tabular}{rrrrrr}
    \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} & \colname{T-Max} \\
    (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
    $0.17$ & $0.18$ & $3.61$ & $0.92$ & $0.88$ & $2.12$ \\
    $0.17$ & $0.18$ & $3.57$ & $0.91$ & $0.88$ & $2.09$ \\
    $0.17$ & $0.18$ & $3.68$ & $0.96$ & $0.88$ & $2.24$ \\
    $0.17$ & $0.17$ & $3.82$ & $0.93$ & $0.9$ & $2.15$ \\
    $0.17$ & $0.17$ & $3.72$ & $0.89$ & $0.9$ & $2.17$ \\
    $0.18$ & $0.18$ & $3.63$ & $0.89$ & $0.9$ & $2.14$ \\
    $0.18$ & $0.16$ & $3.5$ & $0.89$ & $0.89$ & $2.23$ \\
    $0.18$ & $0.17$ & $3.45$ & $0.89$ & $0.93$ & $2.32$ \\
    $0.2$ & $0.17$ & $3.61$ & $0.92$ & $0.88$ & $2.16$ \\
    $0.2$ & $0.18$ & $3.63$ & $0.9$ & $0.92$ & $2.06$ \\
  \end{tabular}

  \begin{tabular}{rrrrrr}
    \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} & \colname{T-Min} \\
    (\colname{SP} \colname{JIT} \colname{SF}) & (\colname{SP} \colname{JIT}) & (\colname{SP}) & (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
    $0.8$ & $0.81$ & $1.3$ & $0.83$ & $0.74$ & $1.2$ \\
    $0.8$ & $0.8$ & $1.28$ & $0.78$ & $0.74$ & $1.22$ \\
    $0.81$ & $0.8$ & $1.31$ & $0.75$ & $0.77$ & $1.22$ \\
    $0.87$ & $0.86$ & $1.29$ & $0.74$ & $0.76$ & $1.18$ \\
    $0.85$ & $0.8$ & $1.26$ & $0.76$ & $0.77$ & $1.19$ \\
    $0.81$ & $0.81$ & $1.3$ & $0.77$ & $0.75$ & $1.22$ \\
    $0.86$ & $0.79$ & $1.25$ & $0.76$ & $0.75$ & $1.17$ \\
    $0.87$ & $0.83$ & $1.24$ & $0.74$ & $0.77$ & $1.16$ \\
    $0.81$ & $0.81$ & $1.3$ & $0.74$ & $0.74$ & $1.17$ \\
    $0.82$ & $0.83$ & $1.23$ & $0.76$ & $0.76$ & $1.16$ \\
  \end{tabular}

  \begin{tabular}{rrr}
    \colname{Orig} & \colname{Orig} & \colname{Orig} \\
    (\colname{JIT} \colname{SF}) & (\colname{JIT}) & () \\\hline
    $0.78$ & $0.75$ & $1.15$ \\
    $0.74$ & $0.73$ & $1.19$ \\
    $0.75$ & $0.72$ & $1.19$ \\
    $0.73$ & $0.73$ & $1.18$ \\
    $0.83$ & $0.76$ & $1.16$ \\
    $0.82$ & $0.72$ & $1.24$ \\
    $0.75$ & $0.72$ & $1.15$ \\
    $0.72$ & $0.76$ & $1.2$ \\
    $0.72$ & $0.72$ & $1.15$ \\
    $0.74$ & $0.73$ & $1.2$ \\
  \end{tabular}
\end{table}

\begin{table}[tp]
  \caption{Exact commands to invoke each benchmark (1/2)}
  \label{t:mb:commands1}
  \tiny\centering

@@

    Fannkuch \colname{T-Min} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_fannkuch\_static\_basic2.txt -X jit-enable-jit-list-wi} \\

@@

    Fannkuch \colname{T-Min} & \texttt{time ./python -X install-strict-loader Tools/benchmarks/fannkuch\_static\_basic2.py 5} \\

@@

    Fannkuch \colname{T-Min} & \texttt{time ./python Tools/benchmarks/fannkuch\_static\_basic2.py 5} \\

@@

@@

\begin{table}[tp]
  \caption{Exact commands to invoke each benchmark (2/2)}
  \label{t:mb:commands2}
  \tiny\centering

@@

    DeltaBlue \colname{T-Min} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_deltablue\_static\_basic2.txt -X jit-enable-jit-list-w} \\

@@

    DeltaBlue \colname{T-Min} & \texttt{time ./python -X install-strict-loader Tools/benchmarks/deltablue\_static\_basic2.py 100} \\

@@

    DeltaBlue \colname{T-Min} & \texttt{time ./python Tools/benchmarks/deltablue\_static\_basic2.py 100} \\

@@

    NBody \colname{T-Max} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_nbody\_static.txt -X jit-enable-jit-list-wildcards -X} \\
    (\colname{SP} \colname{JIT} \colname{SF}) & \texttt{ jit-shadow-frame -X install-strict-loader Tools/benchmarks/nbody\_static.py} \\
    NBody \colname{T-Max} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_nbody\_static.txt -X jit-enable-jit-list-wildcards -X} \\
    (\colname{SP} \colname{JIT}) & \texttt{ install-strict-loader Tools/benchmarks/nbody\_static.py} \\
    NBody \colname{T-Max} & \texttt{time ./python -X install-strict-loader Tools/benchmarks/nbody\_static.py} \\
    NBody \colname{T-Max} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_nbody\_static.txt -X jit-enable-jit-list-wildcards -X} \\
    (\colname{JIT} \colname{SF}) & \texttt{ jit-shadow-frame Tools/benchmarks/nbody\_static.py} \\
    NBody \colname{T-Max} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_nbody\_static.txt -X jit-enable-jit-list-wildcards To} \\
    (\colname{JIT}) & \texttt{ols/benchmarks/nbody\_static.py} \\
    NBody \colname{T-Max} & \texttt{time ./python Tools/benchmarks/nbody\_static.py} \\
    NBody \colname{T-Min} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_nbody\_static\_basic.txt -X jit-enable-jit-list-wildca} \\
    (\colname{SP} \colname{JIT} \colname{SF}) & \texttt{rds -X jit-shadow-frame -X install-strict-loader Tools/benchmarks/nbody\_static\_basic.py} \\
    NBody \colname{T-Min} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_nbody\_static\_basic.txt -X jit-enable-jit-list-wildca} \\
    (\colname{SP} \colname{JIT}) & \texttt{rds -X install-strict-loader Tools/benchmarks/nbody\_static\_basic.py} \\
    NBody \colname{T-Min} & \texttt{time ./python -X install-strict-loader Tools/benchmarks/nbody\_static.py} \\
    NBody \colname{T-Min} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_nbody\_static\_basic.txt -X jit-enable-jit-list-wildca} \\
    (\colname{JIT} \colname{SF}) & \texttt{rds -X jit-shadow-frame Tools/benchmarks/nbody\_static\_basic.py} \\
    NBody \colname{T-Min} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_nbody\_static.txt -X jit-enable-jit-list-wildcards To} \\
    (\colname{JIT}) & \texttt{ols/benchmarks/nbody\_static.py} \\
    NBody \colname{T-Min} & \texttt{time ./python Tools/benchmarks/nbody\_static\_basic.py} \\
    NBody \colname{Orig} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_main.txt -X jit-enable-jit-list-wildcards -X jit-sha} \\
    (\colname{JIT} \colname{SF}) & \texttt{dow-frame Tools/benchmarks/nbody.py} \\
    NBody \colname{Orig} & \texttt{time ./python -X jit -X jit-list-file=Tools/benchmarks/jitlist\_main.txt -X jit-enable-jit-list-wildcards Tools/benc} \\
    (\colname{JIT}) & \texttt{hmarks/nbody.py} \\
    NBody \colname{Orig} & \texttt{time ./python Tools/benchmarks/nbody.py} \\
